### 2015-02-02 paper.tex
create a template for MCM

\documentclass[12pt,a4paper,titlepage]{article}
%\usepackage[doublespacing]{setspace}
\usepackage[utf8]{inputenc}         %This is used for ASKII digits.
\usepackage{amsmath, amssymb}       %This is for mathematic statements,check"amath.colorado.edu/documentation/LaTex/Symbols.pdf"
\usepackage{amsfonts,mathrsfs}      %fonts~~
\usepackage{graphicx}               %This is for graph inserting.
\usepackage{paralist}               %Give me compact lists!!!!! 'compactitem' 'compactenum' 'compactdesc'
%\usepackage{bm}                     %various kinds of bold texts.
\usepackage[top =2.54cm, bottom =2.54cm, left =3.18cm, right =3.18 cm]{geometry}
\usepackage{indentfirst}            %Indent the first letter
\usepackage{fancyhdr}               %Below is the head of the article.
\pagestyle{fancy}
\usepackage{lastpage}
\lhead{Team \# 30061}
\rhead{Page \thepage{} of \pageref{LastPage}}
\cfoot{}
%\numberwithin{equation}{section}    %Numbering of the equations will be based on sections.
%\usepackage{amsthm}
%\theoremstyle{definition}
%\newtheorem{def}{Definition}  %Below set the theorem system.
%\theoremstyle{plain}
%\newtheorem{thm}[law]{Thm}
%\theoremstyle{remark}
%\newtheorem*{remark}{Remark}
\newcommand{\boldit}[1]{\textbf{\textit{#1}}}
\usepackage{float}
\usepackage{paralist}
\usepackage{url}

\begin{document}

\title{}
\date{}
\maketitle

\tableofcontents

\newpage

\section{Introduction}
\label{sec:introduction}

\subsection{Restatement of the Problem}
\label{sec:restatement-of-the-problem}

\subsection{Literature Review}
\label{sec:literature-review}

\begin{itemize}
    \item
    \item
    \item
\end{itemize}

\section{Terminology}
\label{sec:terminology}

\begin{itemize}
    \item
    \item
    \item
\end{itemize}

\section{Assumptions and Justifications}
\label{sec:assumptions-and-justifications}

\begin{itemize}
    \item
    \item
    \item
\end{itemize}

%the main section of the paper
\section{}
\label{sec:}

%some other section of the paper
\section{}
\label{sec:}

\section{Performance and Analysis}
\label{sec:performance-and-analysis}

\subsection{General Performance}
\label{sec:general-performance}

\subsubsection*{Results and Analysis}

%some more specific results and analysis
\subsection{}
\label{sec:}

\subsubsection*{Results and Analysis}

\section{Sensitivity Analysis}
\label{sec:sensitivity-analysis}

%for some improvement of the problem
\section{}
\label{sec:}

\section{Strengths and Weaknesses}
\label{sec:strengths-and-weaknesses}

\subsection*{Strengths}
\label{sec:strengths}

\begin{itemize}
    \item
    \item
    \item
\end{itemize}

\subsection*{Weaknesses}
\label{sec:weaknesses}

\begin{itemize}
    \item
    \item
    \item
\end{itemize}

\section{Conclusions}
\label{sec:conclusions}

%the reference
\begin{thebibliography}{99}

\bibitem{}

\end{thebibliography}

\label{LastPage}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End: